---
title: "EFS for TARGET Landmark paper"
author: "Jenny Smith"
date: "April 20, 2017"
output: pdf_document
---


```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center')
knitr::opts_knit$set(root.dir = '~/COG_Meetings_Analysis/2017.04.18_TARGET_LandmarkPaper/')
```


```{r message = FALSE, warning=FALSE}
library(survival)
library(RColorBrewer)
library(colorspace)
library(stringr)
library(Biobase)
library(dplyr)
library(magrittr)
library(tables)
library(VennDiagram)
getwd()
```


```{r}
source("H:/scripts/Survival_Analysis/Survplot_Functions_2017.03.31.r")
```


# Define Functions
```{r}
#create a list of dataframes from the names of the df objects 
mergeDF <- function(listOfNames){
  #given a listOfNames (character vector) of the dataframes, create a list of these dfs and merge them. 
  
  listOfDF <- list()
  groups <- NULL
  i <- 1
  
  #Create a list with the subsetted dataframes. 
  for (i in 1:length(listOfNames)){
    name <- listOfNames[i]
    
    df <- get(name) 
    listOfDF[[name]] <- df
    
    status <- rep(listOfNames[i], nrow(df)) #make a vector of the mutation status (eg FLT3+,NPM1-)
    groups <- c(groups, status)
  }
  
  #merge the dataframes. 
  merged <- do.call("rbind", listOfDF)
  merged$group <- groups #add column with the mutation status of the patients
  listOfDF[["merged"]] <- merged
  
  # list <- list(listOfDF, merged)
  
  return(listOfDF)
}
```


```{r}
#Checks for the combination of Mutations by looking at unique rows. 
uniqueRows <- function(listOfDF, cols){
  #listOfDF input a named list of dataframes
  #cols is a numeric or character vector of the columns to select
  for (df in names(listOfDF)){
    tmp <- listOfDF[[df]][,cols]
    print(unique(tmp, margin=1))
  }  
}
```



SurvivalPlot is a function that is my generic version in Survplot_Functions_2017.03.31.r
```{r}
efsCustomPlot <- function(fit, diff){
    
    
    #label with pvalue
    pVal <- 1-pchisq(diff$chisq, (length(diff$n)-1))
    pVal <- round(pVal, digits = 4)
    
    if (pVal < 0.001){
      p <- "p < 0.001"
    }else{
      p <- paste("p =", pVal, sep = " ")
    }
    
    
    #X axis position for Num
    # pos <- (max(fit$time)*0.75)

    #customized plot 
    efs_plot <- SurvivalPlot(fit, "Mutation Status", "Years",rainbow(5))

    efs_plot <- efs_plot + scale_x_continuous(limits = c(0,6.5), breaks = seq(0,6.5, 1))
    
    efs_plot <- efs_plot + annotate(geom="text", x=6, y=1.0, label= p, size=6)
    
    #return the ggplot object
    return(efs_plot)
}
```


#Read in the datasets to be analyzed

```{r}
newClassification <- read.csv("H:/COG_Meetings_Analysis/2017.03.27_COG_spring2017/New_stratification_outcomes_updated_2017.03.28.csv", stringsAsFactors = FALSE)

head(newClassification[,c(1,11:18)])
```

```{r}
germanSet <- read.csv("H:/COG_Meetings_Analysis/2017.04.18_TARGET_LandmarkPaper/NUP98-JARID1A_Mutations_ClinData.csv", stringsAsFactors = FALSE)

head(germanSet)
```

```{r}
colnames(germanSet)[12] <- "First.Event"
```


```{r}
AAML2961 <- read.csv("EFS_2961_itd_npm_wt1_nup98_data.csv")

colnames(AAML2961) <- c("time.days", "time.years", "FLT3ITD_NEG", "FLT3ITD_P.NUP98_P.orWT1_P","FLT3ITD_P.NPM1_P", "FLT3ITD_P.NPM1_N.NUP98_N.WT1_N" )
head(AAML2961)
```

```{r}
dim(AAML2961)
```


# Prepare the TARGET Survival Data
```{r}
colnames(newClassification[,50:65])
```


```{r}
mutationsData <- newClassification[, c("TARGET.USI","Progression.Free.Survival","First.Event","Event.Free.Survival.Time.in.Days",  "Reason", "FLT3.ITD.positive.","NPM.mutation","WT1.mutation")]

dim(mutationsData)
```

```{r}
mutationsData$NUP98.NSD1.mutation <- ifelse(grepl("NUP98-NSD1", mutationsData$Reason), "Yes", "No") #not really accurate. Not sure where Rhonda found the "negative" status as confirmed by sequencing or PCR. 
```

```{r}
mutationsData$First.Event <- ifelse(grepl("Censored", mutationsData$First.Event), 0, 1)
mutationsData$Progression.Free.Survival <- ifelse(grepl("No Failure", mutationsData$Progression.Free.Survival), 0, 1)

```

```{r}
dim(mutationsData)
```

```{r}
head(mutationsData)
```

```{r}
lapply(mutationsData[,6:9], function(x) as.factor(x) %>% levels)
```

Some of the clinical data became shifted over by one cell, and that is why there are 5 patients with FLT3.ITD who have numeric information in that column. These patients were not included


```{r}
mutationsData[which(grepl("0.21|0.48|0.7|0.81|2.2", mutationsData$FLT3.ITD.positive.)), ]
```


```{r}
FLT3Negatives <- mutationsData[which(grepl("No", mutationsData$FLT3.ITD.positive.)), ]

dim(FLT3Negatives) #783 patients 
```

This is a check that the numbers are similar to the figure that Rhonda made - she had a total of FLT3-ITD negative patients of 799.
I am findinging 783, so I have 16 missining patients... 

##Define the groups
```{r}
FLT3_p_NPM_p <- mutationsData[which( mutationsData$FLT3.ITD.positive. == "Yes" &  mutationsData$NPM.mutation == "Yes"), ] 

FLT3_n_NPM_p <- mutationsData[which( mutationsData$FLT3.ITD.positive. =="No" &  mutationsData$NPM.mutation == "Yes"), ]

FLT3_n_NPM_n <- mutationsData[which( mutationsData$FLT3.ITD.positive. == "No" &  mutationsData$NPM.mutation == "No"), ] 

FLT3_p_NPM_n_wt1_n_NUP98_n <- mutationsData[which( mutationsData$FLT3.ITD.positive. == "Yes" &  mutationsData$NPM.mutation == "No" &  mutationsData$WT1.mutation == "No" &  mutationsData$NUP98.NSD1.mutation == "No"), ]

FLT3_p_wt1_p_or_NUP98_p <- mutationsData[which(mutationsData$FLT3.ITD.positive. == "Yes" & (mutationsData$WT1.mutation == "Yes" | mutationsData$NUP98.NSD1.mutation == "Yes")), ]
```


##Make a list of the of the subsetted dataframes and merge them 
```{r}
targetGroups <- ls(pattern = "FLT3_")

target <- mergeDF(targetGroups)

summary(target)
```

```{r}
sapply(target[1:5], FUN=dim)

```



```{r}
uniqueRows(target[1:5], c(6:9))
```


#Check overlaps of patients in cohorts here





#Prepare the German Set Survival Data

Must convert the numeric codes to yes/no and fill in the time with the time from diagnosis to last followup for patients who are censored. 
```{r}
germanSet$FLT3.ITD <- ifelse(germanSet$FLT3.ITD == 2 | germanSet$FLT3.ITD == 4, "Yes", "No")

germanSet$mutWT1 <- ifelse(germanSet$mutWT1 == 4 | germanSet$mutWT1 == 5 | germanSet$mutWT1 == 6, "Yes", "No")

germanSet$mutNPM1 <- ifelse(germanSet$mutNPM1 == 13, "Yes", "No")

germanSet$NUP98.NSD1 <- ifelse(germanSet$NUP98.NSD1 == 9, "Yes", "No")

germanSet$First.Event <- ifelse(germanSet$First.Event == 0, 0, 1)

germanSet$Time.dx.event1..months. <- ifelse(is.na(germanSet$Time.dx.event1..months.), germanSet$Time.DxtoLastFollowUp.Months, germanSet$Time.dx.event1..months.)
```


```{r}
head(germanSet)
```



#Define the Groups for Survival Analysis
```{r}
length(which(germanSet$FLT3.ITD == "Yes")) #54 positive FLT3ITD
length(which(germanSet$FLT3.ITD == "No")) #279 negative FLT3ITD
```


```{r}
FLT3.p.NPM.p <- germanSet[which(germanSet$FLT3.ITD == "Yes" & germanSet$mutNPM1 == "Yes"), ]

FLT3.n.NPM.p <- germanSet[which(germanSet$FLT3.ITD == "No" & germanSet$mutNPM1 == "Yes"), ]

FLT3.n.NPM.n <- germanSet[which(germanSet$FLT3.ITD == "No" & germanSet$mutNPM1 == "No"), ]

FLT3.p.NPM.n.wt1.n.NUP98.n <- germanSet[which(germanSet$FLT3.ITD == "Yes" & germanSet$mutNPM1 == "No" & germanSet$mutWT1 == "No" & germanSet$NUP98.NSD1 == "No"), ]

FLT3.p.wt1.p.or.NUP98.p <- germanSet[which(germanSet$FLT3.ITD == "Yes" & (germanSet$mutWT1 == "Yes" | germanSet$NUP98.NSD1 == "Yes")),]
```


```{r}
GermanGroupNames <- ls(pattern = "FLT3\\.")

german <- mergeDF(GermanGroupNames)

summary(german)
```

```{r}
sapply(german[1:5], dim)
```

```{r}
uniqueRows(german[1:5], c(7:10))
```


#Survival Analysis of TARGET AML 

```{r}
colnames(target$merged)
```

###Log-Rank test between all 5 curves
```{r}
efs <- SurvObjects(target$merged, c("Event.Free.Survival.Time.in.Days", "First.Event"), "group", "DtoY")

efs
```

###log-rank test between FLT3+,NPM1+ versus FLT3+,NPM1-

```{r}
colnames(target$merged)
```


```{r}
NPM1 <- target$merged[which(grepl("FLT3_p_NPM_p|FLT3_p_NPM_n_wt1_n_NUP98_n", target$merged$group)), ]

unique(NPM1$group)
```

```{r}
NPM1_SA <- SurvObjects(NPM1, c("Event.Free.Survival.Time.in.Days", "First.Event"), "group", "DtoY")
NPM1_SA
```


###FLT3+,NPM1+ and FLT3-,NPM1+ 
```{r}
NPM2 <- target$merged[which(grepl("FLT3_p_NPM_p|FLT3_n_NPM_n", target$merged$group)), ]

unique(NPM2$group)
```


```{r}
NPM2 <- SurvObjects(NPM2, c("Event.Free.Survival.Time.in.Days", "First.Event"), "group", "DtoY" )

NPM2
```



#Survival Analysis with German Data

```{r}
colnames(german$merged)
```


```{r}
efs_Ger <- SurvObjects(german$merged, c("Time.dx.event1..months.", "First.Event"), "group", "MtoY")

efs_Ger
```

```{r}
NPM1gc <- german$merged[which(grepl("FLT3.p.NPM.p|FLT3.p.NPM.n.wt1.n.NUP98.n", german$merged$group)), ] 
NPM2gc <-  german$merged[which(grepl("FLT3.p.NPM.p|FLT3.n.NPM.n", german$merged$group)), ]

unique(NPM1gc$group)
unique(NPM2gc$group)
```


```{r}
NPM1gc_SA <- SurvObjects(NPM1gc, c("Time.dx.event1..months.", "First.Event"), "group", "MtoY")

NPM1gc_SA
```

```{r}
NPM2gc_SA <- SurvObjects(NPM2gc, c("Time.dx.event1..months.", "First.Event"), "group", "MtoY")
NPM2gc_SA
```



#Custom Plots for Publication

##TARGET 0531
```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=11}
efs0531 <- efsCustomPlot(efs[[1]],efs[[2]])

efs0531$labels$title <- "EFS in TARGET AAML0531 Cohort"

efs0531 <- efs0531 + 
  scale_linetype_manual(values = c("twodash", "solid", "longdash", "dashed", "dotdash"), guide=FALSE) +
  scale_color_manual(values = c("FLT3_n_NPM_p"="red", 
                                "FLT3_p_NPM_p"="green4", 
                                "FLT3_n_NPM_n"="darkorchid2", 
                                "FLT3_p_NPM_n_wt1_n_NUP98_n"="blue2",
                                "FLT3_p_wt1_p_or_NUP98_p"="gold3"),
                     labels=c("FLT3ITD-, NPM1+ (N=44)",
                              "FLT3ITD+, NPM1+ (N=34)", 
                              "FLT3ITD-, NPM1- (N=730)", 
                              "FLT3ITD+, NPM1-, WT1-, NUP98:NSD1- (N=77)",
                              "FLT3ITD+, WT1+ or NUP98:NSD1+ (N=57)"), name=" ")


# pdf(file="TARGET_AML_EFS_0531_21April2017.pdf", height = 5, width = 11)
efs0531 
# dev.off()
```

##German Cohort

```{r}
names(german[1:5])
```


```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=11}
efs_german <- efsCustomPlot(efs_Ger[[1]], efs_Ger[[2]])


efs_german$labels$title <- "EFS in German Cohort"

efs_german <- efs_german + 
  scale_linetype_manual(values = c("solid", "twodash", "longdash", "dashed", "dotdash"), guide=FALSE) +
  scale_color_manual(values =c("FLT3.p.NPM.p"="green4", 
                               "FLT3.n.NPM.p"="red", 
                               "FLT3.n.NPM.n"="darkorchid2", 
                               "FLT3.p.NPM.n.wt1.n.NUP98.n"="blue2", 
                               "FLT3.p.wt1.p.or.NUP98.p"="gold3"), 
                     labels=c("FLT3ITD+, NPM1+ (N=8)", 
                              "FLT3ITD-, NPM1+ (N=11)", 
                              "FLT3ITD-, NPM1- (N=268)", 
                              "FLT3ITD+, NPM1-, WT1-, NUP98:NSD1- (N=26)", 
                              "FLT3ITD+, WT1+ or NUP98:NSD1+ (N=20)"), name=" ") 

# pdf(file="EFS_german_21April2017.pdf", height = 5, width = 11)
efs_german
# dev.off()
```



##AAML2961
```{r}
colnames(AAML2961)
```


```{r}
efs_2961 <- ggplot(AAML2961) + 
  geom_line(mapping = aes(x=AAML2961$time.years, y=AAML2961$FLT3ITD_P.NUP98_P.orWT1_P, colour="FLT3ITD+, WT1+ or NUP98:NSD1+ (N=24)", linetype="solid")) +
  geom_line(mapping = aes(x=AAML2961$time.years, y=AAML2961$FLT3ITD_P.NPM1_N.NUP98_N.WT1_N, colour="FLT3ITD+, NPM1-, WT1-, NUP98:NSD1- (N=14)", linetype="dashed")) +
  geom_line(mapping = aes(x=AAML2961$time.years, y=AAML2961$FLT3ITD_P.NPM1_P, colour="FLT3ITD+, NPM1+ (N=7)", linetype="dotdash")) +
  geom_line(mapping = aes(x=AAML2961$time.years, y=AAML2961$FLT3ITD_NEG, colour="FLT3ITD- (N=536)", linetype="longdash")) +
  scale_y_continuous(limits = c(0,1.0), breaks = seq(0,1.0, 0.2)) + 
  scale_x_continuous(limits = c(0,6.5), breaks = seq(0,6.5,1)) + 
  labs(y= "Fraction Surviving", x ="Follow-up in Years") + 
  theme(plot.title = element_text(hjust = 0.5, size=18), 
          panel.background = element_rect(fill="white"), 
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text.y = element_text(size = 14),
          axis.text.x = element_text(size=14), 
          axis.title = element_text(size = 16),
          legend.text = element_text(size=14),
          panel.border = element_rect(colour = "dark grey", fill= NA, size=1.0)) 
```


```{r warning=FALSE, message=FALSE, fig.height=5, fig.width=11}
efs_2961 <- efs_2961 + 
  scale_linetype_manual(values = c("dashed","solid", "longdash", "dotdash"), guide=FALSE) + 
  scale_color_manual(name=" ", values = c( "FLT3ITD+, NPM1+ (N=7)"="green4", 
                                           "FLT3ITD- (N=536)"="darkorchid2", 
                                           "FLT3ITD+, NPM1-, WT1-, NUP98:NSD1- (N=14)"="blue2",
                                           "FLT3ITD+, WT1+ or NUP98:NSD1+ (N=24)"="gold3")) +
  annotate(geom="text", x=6, y=1.0, label= "p < 0.001", size=6) +
  labs(title="EFS in TARGET AAML2961 Cohort")

# pdf(file="TARGET_AML_EFS_2961_21April2017.pdf", height = 5, width = 11)
efs_2961
# dev.off()
```



